---
title: "Graphs Describing the Damage From Various Extreme Weather Events"
---

In this analysis I used storm data put out by the National Weather Service to find the extent of the impacts of extreme weather events since 2000, in terms of casualties, deaths, and property/crop damage.

#Data Processing
Load the raw data into R
```{r, cache=TRUE}
if(!exists("rawdata"))
    read.csv("repdata_data_StormData.csv.bz2") -> rawdata
```

Store the raw data into a data frame which will be manipulated
```{r, cache=TRUE}
rawdata->stormData
```

Getting the date
```{r, cache=TRUE}
as.POSIXct(strptime(stormData$BGN_DATE, "%m/%d/%Y %H:%M:%S"))->stormData$BGN_DATE
```

Removing all dates before 2000
```{r, cache=TRUE}
stormData[stormData$BGN_DATE>as.POSIXct("2000-1-1"), 1:37] -> stormData
```

Fixing damage values
```{r, cache=TRUE}

stormData->temps
gsub("^$", 0,temps$CROPDMGEXP)->temps$CROPDMGEXP
gsub("^[kK]$", 1,temps$CROPDMGEXP)->temps$CROPDMGEXP
gsub("^[mM]$", 1000,temps$CROPDMGEXP)->temps$CROPDMGEXP
gsub("^[bB]$", 1000000,temps$CROPDMGEXP)->temps$CROPDMGEXP
gsub("^[?02]$", 0,temps$CROPDMGEXP)->temps$CROPDMGEXP

gsub("^[?0+-23456789]$", 0,temps$PROPDMGEXP)->temps$PROPDMGEXP
gsub("^$", 0,temps$PROPDMGEXP)->temps$PROPDMGEXP
gsub("^[kK]$", 1,temps$PROPDMGEXP)->temps$PROPDMGEXP
gsub("^[mM]$", 1000,temps$PROPDMGEXP)->temps$PROPDMGEXP
gsub("^[bB]$", 1000000,temps$PROPDMGEXP)->temps$PROPDMGEXP
gsub("^[hH]$", .1,temps$PROPDMGEXP)->temps$PROPDMGEXP

as.numeric(temps$PROPDMGEXP)*temps$PROPDMG->temps$PROPDMG
as.numeric(temps$CROPDMGEXP)*temps$CROPDMG->temps$CROPDMG
temps$PROPDMG+temps$CROPDMG->temps$PROPDMG

temps->stormData
```

Getting a General list of events with more than 50 instances (to weed out anomalies)
```{r, cache=TRUE}
as.data.frame(table(stormData$EVTYPE))->stormTypes
stormTypes[stormTypes$Freq>1, 1:2]->stormTypes
```

```{r}
library(dplyr)
```

523149 before this operation
```{r, cache=TRUE}
stormData[stormData$EVTYPE%in%stormTypes$Var1, 1:37] -> sStorms
```



```{r}
library(dplyr)
sStorms$EVTYPE->sStorms$Type
sStorms %>%
    group_by(Type) %>% #Groups the data so it can be processed by summarise
    summarise(Damage=sum(PROPDMG))->propdamage #finds the sum of emissions by year-type pair
```

```{r}
propdamage[propdamage$Damage>2000000, 1:2] ->propdamage
```

```{r}
library(dplyr)
sStorms$EVTYPE->sStorms$Type
sStorms %>%
    group_by(Type) %>% #Groups the data so it can be processed by summarise
    summarise(Injuries=sum(INJURIES), Fatalities=sum(FATALITIES))->Casualties #finds the sum of emissions by year-type pair
```

```{r}
Casualties[Casualties$Injuries>315, 1:3] ->Casualties
```

```{r}
library(lattice)
with(propdamage, barchart(Type~Damage,propdamage,main = "Damage from Various Types of Storms from 2000 to 2011", xlab = "Damage in Thousands of Dollars"))
```

```{r}
library(lattice)
with(Casualties, barchart(Type~Injuries,Casualties,main = "Injuries from Various Types of Storms from 2000 to 2011", xlab = "Injuries", xlim = c(0,15500)))
```
```{r}
library(lattice)
with(Casualties, barchart(Type~Fatalities,Casualties,main = "Fatalities from Various Types of Storms from 2000 to 2011", xlab = "Injuries", xlim = c(0,1300)))
```
